\section*{Acknowledgements}
It’s with deep gratitude that I pick up the pen today to express my sincere thanks to all those who have contributed to the realization of this Master's Engineering thesis. This work represents the culmination of several years of effort and learning, and I am honored to share this moment with you.

Firstly, I would like to thank Dr. Marc TEYSSIER and Dr. Clément DUHART for their invaluable guidance, constant support, and insightful advice. Your expertise and passion for research have been a source of inspiration.

My thanks also go to the entire faculty of the DVIC, including the PI and doctoral students. Your high-quality teaching and willingness to share knowledge have enriched my academic and professional journey.

I want to express my gratitude to my colleagues from DVIC for their teamwork, exchanges, and mutual support. Our discussions and debates have shaped my understanding of the topics covered in this thesis, and I am honored to have shared this experience with all of you.

Finally, I would like to thank Violette ANICET-MERVEILLEUX for her help with the bio-patches and bio-materials for this project. And I would like to express my gratitude to all those who have contributed, directly or indirectly, to completing this research work. Your advice, feedback, and contributions have been invaluable in accomplishing this thesis.

Finally, I would like to express my deep appreciation to my friends and family for their continuous support and positivity during the years of my studies.
