\chapter{First Step into the Field of Haptic}
\section{Introduction}

\section{Related Work}

\section{Evaluation}
The device has been tested with various sensors for different applications. A typical pattern repeats: 

\subsection{Setup}
A sensor is used and placed at the user's fingertip. It is connected to a micro-controller, in this case, a Teensy. The Teensy is linked to the DRV2605 haptic controller, which controls the haptic motors. The LRA or ERM motors must be in contact with the user's skin. Choosing a relatively sensitive skin area is preferable to perceiving the vibrations better.

\subsection{Protocol}
Connect the Teensy to a power source. Launch the code corresponding to the sensor used. The software allows real-time visualization of the recorded data, which helps the user better understand the differences between the vibrations they should feel during the test.

\textit{Simple Piezo Sensor}
To measure the pressure exerted by the user with their fingers. Perform various object manipulations, handling them gently or applying more force, pushing lift different objects. Regarding texture, slide the Piezo sensor on other surfaces; it will generate additional data based on surface variations.

\textit{Flexible Piezo Sensor}
To evaluate the surface of objects and the user's environment. Slide your fingers along surfaces; as soon as they come into contact, the Piezo sensor bends and records data transmitted to the rest of the device.

\textit{Sound Sensor}
Place the sensor at the tip of your fingers and run them over different textures and surfaces. The sound your fingers produce on the material is recorded and transmitted to the controller.

The sensor can be considered functional if the variations in vibrations correspond to the different actions of the user and the displayed data. Currently, the sensors that have shown the most efficiency are the Flexible Piezo types.

\section{Limitations}
It’s essential to note that this chapter serves as an introduction and may not necessarily represent a significant contribution for individuals with advanced expertise in this field. It aims to remain accessible to a broad audience, whether in innovation and technology, design and engineering, or even economic aspects. That's why it delves deeply into haptic to ensure clear understanding while keeping it simple.

\textit{Hardware}

Currently, the prototype uses only one sensor connected to the micro-controller. The next phase, which involves placing a sensor on each finger of the prosthesis, could pose constraints in power, data reading capacity, and data output. The DRV2605 haptic motor driver controls only one ERM or LRA at a time. The overall size of the components can be excessive, limiting its use for certain prostheses and users. Additionally, the motors used are currently too bulky. The haptic effect becomes much less interesting by reducing the mass and increasing the frequency.

\textit{software}

The current code continuously reads data in real time, which keeps it in constant operation. The accumulation of data and some noise can slow down the entire system.

\section{Future Works}
Innovation in Haptic is still relatively new and represents a path for improvement in engineering. As mentioned earlier, the work presented in this chapter marks a first step and opens the way for numerous potential enhancements.

From a hardware perspective, one improvement for this project is creating a printed circuit board (PCB) to reduce the space occupied by all the components. This is crucial to ensure the efficient use of prosthetics, ensuring that the entire setup is optimal in terms of distance without causing interference or adding extra constraints to the user. Given the rapid evolution in Haptic, it's also essential to stay updated on new components to determine if they might be better suited and more efficient for this project. Using a controller capable of driving multiple DRV2605 devices is worth considering.

There are opportunities to enhance the project on the software front by integrating new features related to the components into the code. By optimizing the code for faster and more efficient operation, it becomes possible to improve the precision of vibrations. Other features associated with the controllers can be explored to optimize the code and offer new functionalities.

\section{Conclusion}
This chapter represents a first step towards understanding and accessing haptic technology. It explores various types of components, their functionalities, and different ways to use them. Furthermore, it offers multiple practical applications and explicit code that everyone can use. It aims to remain simple and accessible to all. The next chapter delves into prosthetics and presents different prototypes for integrating the haptic components introduced in this chapter.